Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2021-12-06 Manuscript/Manuscript.Rmd
Use correct icc function, add meeting comments

# plot
miss_ind %>% 
  mutate(name = factor(name, levels = unique(name))) %>% 
ggplot(aes(
  x = name,
  y = rownr,
  color = value,
  width = 0.8, 
  height = 0.8,
  label = text
)) +
  geom_tile(fill = "white", size = 2) +
  geom_text(show.legend = FALSE) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(labels = 5:1) +
  scale_color_manual(values = plot_col, name = "") +
  theme_minimal() +
  theme(
    legend.position = "bottom", 
    panel.grid.major = element_blank()) +
  labs(x = "", y = "")
# TODO: make it one plot with sporadic and systematic
# TODO: participant/response/*sample units*/records instead of obs on the y axis

@@

# ICCs
icc_pop <- multilevel::ICC1(aov(popular ~ as.factor(class), data = pop))
icc_teach <- multilevel::ICC1(aov(popteach ~ as.factor(class), data = pop))
```

@@

# x = "popular"
# z = "popteach"
plot_conditional <- function(data, x, z, cluster, ...){
  ggplot(data, aes(x = get(x), color = factor(is.na(get(z)), labels = c("missing", "observed")))) +
  geom_density(data = data %>% filter(!is.na(get(z))), aes(x = get(x), group = get(cluster)), alpha = 0.01, fill = plot_col[1], color = NA) +
  geom_density(data = data %>% filter(is.na(get(z))), aes(x = get(x), group = get(cluster)), alpha = 0.01, fill = plot_col[2], color = NA) +
  geom_density() +
  scale_color_manual(values = plot_col, name = z) +
  theme_classic() +
  theme(legend.position = "bottom") +
  labs(x = x)
}

@@

@@

# compare ICCs before and after imputation
ICCs <- data.frame(
  vars = c("popular", "popteach", "texp"), 
  incomplete = c(multilevel::ICC1(aov(popular ~ class, pop)), 
               multilevel::ICC1(aov(popteach ~ class, pop)),
               multilevel::ICC1(aov(texp ~ class, pop))), 
  ignored = c(multilevel::ICC1(aov(popular ~ class, complete(imp_ignored))), 
              multilevel::ICC1(aov(popteach ~ class, complete(imp_ignored))), 
              multilevel::ICC1(aov(texp ~ class, complete(imp_ignored))))
  )
ICCs
```

@@

```{r pop_predictor, echo=FALSE, message=FALSE, warning=FALSE}
# adjust the predictor matrix
pred <- ini$pred 
pred[, "pupil"] <- 0

@@

# compare ICCs before and after imputation
ICCs <- ICCs %>% cbind(
           predictor = c(multilevel::ICC1(aov(popular ~ class, complete(imp_predictor))), 
                        multilevel::ICC1(aov(popteach ~ class, complete(imp_predictor))), 
                        multilevel::ICC1(aov(texp ~ class, complete(imp_predictor))))
           )
ICCs
```

@@

>>>>

# plot
miss_ind %>% 
  mutate(name = factor(name, levels = unique(name))) %>% 
ggplot(aes(
  x = name,
  y = rownr,
  color = value,
  width = 0.8, 
  height = 0.8,
  label = text
)) +
  geom_tile(fill = "white", size = 2) +
  geom_text(show.legend = FALSE) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(labels = 5:1) +
  scale_color_manual(values = plot_col, name = "") +
  theme_minimal() +
  theme(
    legend.position = "bottom", 
    panel.grid.major = element_blank()) +
  labs(x = "", y = "")
# TODO: make it one plot with sporadic and systematic
# TODO: participant/response/*sample units*/records instead of obs on the y axis
# TODO: cluster 3
# TODO: add border/line
# TODO: remove unit
# TODO: add accolade

@@

# ICCs
icc_pop <- icc(popular ~ as.factor(class), data = pop)
icc_teach <- icc(popteach ~ as.factor(class), data = pop)
```

@@

# x = "popular"
# z = "popteach"
plot_conditional <- function(data, x, z, cluster, ...){
  ggplot(data, aes(x = get(x), color = factor(is.na(get(z)), labels = c("missing", "observed")))) +
  geom_density(data = data %>% filter(!is.na(get(z))), aes(x = get(x), group = get(cluster)), alpha = 0.01, fill = plot_col[2], color = NA) +
  geom_density(data = data %>% filter(is.na(get(z))), aes(x = get(x), group = get(cluster)), alpha = 0.01, fill = plot_col[1], color = NA) +
  geom_density() +
  scale_color_manual(values = plot_col, name = z) +
  theme_classic() +
  theme(legend.position = "bottom") +
  labs(x = x)
}

@@

# marginal

@@

# compare ICCs before and after imputation
ICCs <- data.frame(
  vars = c("popular", "popteach", "texp"), 
  incomplete = c(icc(popular ~ class, pop), 
               icc(popteach ~ class, pop),
               icc(texp ~ class, pop)), 
  ignored = c(icc(popular ~ class, complete(imp_ignored)), 
              icc(popteach ~ class, complete(imp_ignored)), 
              icc(texp ~ class, complete(imp_ignored)))
  )
ICCs
```

@@

```{r pop_predictor, message=FALSE, warning=FALSE}
# adjust the predictor matrix
pred <- ini$pred 
pred[, "pupil"] <- 0

@@

# compare ICCs before and after imputation
ICCs <- ICCs %>% cbind(
           predictor = c(icc(popular ~ class, complete(imp_predictor)), 
                        icc(popteach ~ class, complete(imp_predictor)), 
                        icc(texp ~ class, complete(imp_predictor)))
           )
ICCs
```

@@

- Adding a ICC=0 dataset to show that even if there is no clustering it doesn't hurt.

### 2021-12-06 Manuscript/Manuscript.tex
Use correct icc function, add meeting comments

\begin{CodeOutput}
      vars incomplete   ignored
1  popular  0.3280070 0.2958571
2 popteach  0.3138658 0.2561198
3     texp  1.0000000 0.4696024
\end{CodeOutput}
\end{CodeChunk}

@@

@@

\begin{CodeOutput}
      vars incomplete   ignored predictor
1  popular  0.3280070 0.2958571 0.3581435
2 popteach  0.3138658 0.2561198 0.3446899
3     texp  1.0000000 0.4696024 1.0000000
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{itemize}
\item
  Adding some kind of help function to mice that suggests a suitable
  predictor matrix to the user, given a certain analysis model.
\item
  Adding a \texttt{multilevel\_ampute()} wrapper function in mice.
\item
  Exporting \texttt{mids} objects to other packages like \texttt{lme4}
  or \texttt{coxme}?
\end{itemize}

>>>>

\begin{CodeOutput}
      vars incomplete   ignored
1  popular  0.3280070 0.2756762
2 popteach  0.3138658 0.2456775
3     texp  1.0000000 0.4129275
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # adjust the predictor matrix
R> pred <- ini$pred 
R> pred[, "pupil"] <- 0
R> 
R> # impute the data, cluster as predictor
R> imp_predictor <- mice(pop, maxit = 10, pred = pred, print = FALSE)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeOutput}
      vars incomplete   ignored predictor
1  popular  0.3280070 0.2756762 0.3587923
2 popteach  0.3138658 0.2456775 0.1998760
3     texp  1.0000000 0.4129275 1.0000000
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{itemize}
\item
  Adding some kind of help function to mice that suggests a suitable
  predictor matrix to the user, given a certain analysis model.
\item
  Adding a \texttt{multilevel\_ampute()} wrapper function in mice.
\item
  Exporting \texttt{mids} objects to other packages like \texttt{lme4}
  or \texttt{coxme}?
\item
  Adding a ICC=0 dataset to show that even if there is no clustering it
  doesn't hurt.
\end{itemize}
